\section{Expected Outcomes and Budget}
\label{sec:outcomes}

\subsection{Expected Outcomes}

\subsubsection{Operational Reliability}

\textbf{Telemetry Performance:} Continuous data stream capability maintaining operational integrity up to 5 km altitude with verified packet loss rate below 5\% utilizing XBee-PRO 900HP telemetry system. Real-time transmission of GPS position data enabling accurate vehicle tracking throughout flight and recovery phases.

\textbf{Safe Recovery Operations:} Successful parachute deployment achieving zero thermal damage to canopy materials through implementation of sealed explosion cap design and PWM-controlled ignition system. Validated deployment consistency eliminating single-use deployment limitations.

\subsubsection{System Integrity}

\textbf{Electronics Reliability:} Custom-fabricated PCB architecture eliminating wire fatigue failure modes while ensuring reliable electrical connections under sustained vibration conditions throughout launch and flight phases.

\textbf{Mechanical Robustness:} Vibration-proof 3D-printed avionics holder maintaining center of gravity stability and protecting sensitive electronic components from high-G launch forces and deployment shock loads.

\subsubsection{Ground Station Capabilities}

\textbf{Real-time Monitoring:} Comprehensive ground station interface providing continuous monitoring of critical parameters including battery voltage, igniter continuity status, and GPS position data.

\textbf{Safety Protocol Implementation:} Automatic "Launch Inhibit" safety system preventing launch operations when main battery voltage drops below 14.0V threshold, eliminating deployment failures due to insufficient power conditions.

\subsection{Project Timeline}

The project implementation follows a structured timeline ensuring systematic development and validation of all subsystems:

\begin{table}[h]
\centering
\caption{Project Timeline and Milestones}
\label{tab:timeline}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Phase} & \textbf{Duration} & \textbf{Key Activities} \\ \midrule
Literature Review & Weeks 1-2 & Comprehensive review of systems \\
System Design & Weeks 3-5 & PCB design, CAD modeling \\
Component Procurement & Weeks 4-6 & Purchase components \\
PCB Fabrication & Weeks 6-7 & PCB manufacturing \\
Mechanical Fabrication & Weeks 7-9 & 3D printing assembly \\
System Integration & Weeks 9-10 & Integration of subsystems \\
Ground Testing & Weeks 10-12 & Static tests, range tests \\
System Validation & Weeks 12-13 & Full system validation \\
Flight Preparation & Week 14 & Launch readiness \\ \bottomrule
\end{tabular}
\end{table}

\subsection{Budget Considerations}

The project budget encompasses electronic components, mechanical fabrication materials, and testing resources. Items marked with asterisk (*) indicate components already purchased and available, reducing overall project cost.

\subsubsection{Major Component Costs}

\begin{itemize}
    \item XBee-PRO 900HP Modules (2 units): Approximately KES 15,000
    \item ESP32 Development Boards: KES 2,000
    \item Custom PCB Fabrication: KES 5,000
    \item LiPo Batteries and BMS: KES 8,000*
    \item 3D Printing Materials: KES 3,000
    \item Nichrome Wire and Igniters: KES 2,000
    \item Electronic Components (MOSFETs, passives, connectors): KES 4,000
    \item Testing Equipment Access: Available through university facilities
\end{itemize}

Total estimated budget: Approximately KES 39,000 (with some components already available).
